---
title: "Exercise 03 - Practice Options Data (without answers)"
output: html_document
---


This exercise set is going to be structured a little been differently.  In the first place, all the questions are in this R Markdown file, because I want you to get comfortable with this file format.  Additionally, many of the questions will start with a completed query that you will be asked to modify.  Much can be learned from reading completed code and understanding it enough to modify it.

This document is not intended to be `knit`.  Instead, read through it in the RStudio source window, and execute the code in chunks (ctrl/cmd + shift + enter) or by highlighting lines (ctrl/cmd + enter).  I have been using in-line output as I create this document and I'm enjoying.  I encourage you to expriment with both in-line and console output and see what you like best.  

&nbsp;
&nbsp;

#### Preliminaries

1. Check the working directory of this file and then try to change the working directory to your desktop.  To run all a single line of code in a chunk press `ctrl/cmd + enter`, and the line of code that your cursor is on will be executed.  To run an entire chunk, press `ctrl/cmd + shift + enter`.  You don't have to knit the document to run code.

```{r}
# You should get an error when you try to change the working directory.  Basically, R Markdown notebooks are setup to keep the working directory the same as the folder that contains that the .Rmd file itself.

```

&nbsp;

2. Load `readr` and `dplyr` into this R session.
```{r}
library(readr)
library(dplyr)

```

&nbsp;

3. Put the two CSVs `data_practice_market_history` and `data_practice_option_history` into the same folder as this .Rmd file and then read in the CSVs (using `readr::read_csv()`) into two data frames called `df_market_history` and `df_option_history`.  
```{r}
# When you use `read_csv()` put the following code as the second arugument `col_types = cols()` in order to avoid annoying warning messages.
df_market_history <- read_csv("data_practice_market_history.csv", col_types = cols())
df_option_history <- read_csv("data_practice_option_history.csv", col_types = cols())
```

&nbsp;
&nbsp;

#### Exploring `df_market_history`

4. Let's look at all the underlyings that are represented in `df_market_history` on 8/16/2018.  How many are there?  Modify the query to see how many underlyings there are on 12/16/2013?
```{r}
df_market_history %>% 
  filter(trade_date == as.Date("2013-08-16")) %>% 
  distinct(underlying)

```

&nbsp;

5. The following query counts the number of underlyings that are accounted for on each trade date in `df_market_history`.  About how many underlyings are there each day?
```{r}
df_market_history %>% 
  group_by(trade_date) %>% 
  summarize(num_und = n_distinct(underlying)) %>% 
  arrange(trade_date)


```

&nbsp;


6. The following query returns a list of all the trade dates that are represented in `df_market_history`. Add a use of `top_n()` with `%>%` in order to isolate the earliest date and the latest date.
```{r}
# earliest is 8/16/2013
# latest is 1/16/2014
df_market_history %>% distinct(trade_date) 

```

&nbsp;

7. The follow query returns a tibble consisting of all the distinct expirations represented in the data for a trade date of 8/16/2013 - there should be only one, 9/21/2013.  Modify the query to see how many SPY expirations are in the data set for a trade date of 9/20/2013.  Is it the same expiration that you found for 8/16?  See what happens when you remove the `underlying == "SPY"` filter - what conclusion can you make?
```{r}
df_market_history %>% 
  filter(underlying == "SPY") %>% 
  filter(trade_date == as.Date("2013-08-16")) %>% 
  distinct(expiration)


```

&nbsp;


8.  The follow query counts the number of distinct `expiration` that can be found for each `trade_date`. What I want you to see is that for each `trade_date`, there is only one `expiration` in the data set.  There 106 trade dates, so we could really just look at the entire dataframe by piping into `View()` to verify that `num_exps = 1` for all trade dates.  However, if there were thousands of trade dates this would be less feasible.  

Modify this code so that you can check the `min()` and `max()` of `num_exps`.  An extra 100 nerd points if you can do it without creating an intermediate dataframe - i.e. if you can do it with just pipes.

```{r}
# The main point of the exercise is just to see that for each trade date there is only data for one expiration.
df_market_history %>% 
  group_by(trade_date) %>% 
  summarize(num_exps = n_distinct(expiration)) %>% 
  arrange(trade_date)
                      
                      
```


&nbsp;


9. The following query shows which `expiration` is given on each `trade_date`.  Why is it OK that I use `min()` in the summarize function?  Try using `max()` instead and see that you get the same answer.  Pipe this query into `View()` and look at how `exp_date` changes as you move through time.
```{r}
df_market_history %>% 
  group_by(trade_date) %>% 
  summarize(exp_date = min(expiration)) %>% 
  arrange(trade_date)

```


&nbsp;


10.  The following query checks for OTM (out-of-the-money) calls in `df_market_history` - notice that we use `implied_forward` to determine moniness of options.  Modify the query to check for ITM (in-the-money) calls.  Did you find any?  Create another query to check for OTM/ITM puts.  Do you find any ITM puts?

```{r}
df_market_history %>% filter(type == "call") %>% 
                      filter(strike > implied_forward)

```

&nbsp;

**Summary**

Here are some facts about the `df_market_history` dataset.  We have demonstrated a lot of these by the previous exercises.  Note that the numbers below do not correspond to the exercises above.

1. **Context:** For nearly all underlyings, there is one expiration date per month - the third Saturday.  For most underlyings, 3 or 4 expirations are actively traded at any given time.  The nearest expiration is sometimes referred to as the *front-month*.

2. In `df_market_history`, for each trade date, only front-month options are included.

3. Notice that on Friday 8/16/2013, all the options have an expiration date of 9/21/2013.  This remains true until Friday 9/20/2013, which is the final trading day of the 9/21 expiration.  On that day, the all options have an expiration of 10/19/2013, which remains the case until Friday 10/18.

4. This data set contains option data from the 100 most liquid ETF underlyings.

5. Only OTM options are represented in the data.

6. **Putting It Together:** `df_market_history` consists of EOD option prices from 8/16/2013 to 1/16/2014, for about 100 liquid ETF underlyings.  On any given day, only fronth-month, OTM option prices are included.

7. **Looking Ahead:** for the option trading strategies we are going to consider, we will only be trading front-month, OTM options.


&nbsp;
&nbsp;

#### Exploring `df_option_history`

11. We can use the following query see all the dates for which the September 2013 SPY 166 put was an OTM option. Recall that `df_market_history` starts at 8/16/2013, so we can conclude that this options was ITM from 8/16 to 8/22.  Also, notice the gap between 8/23 and 9/9, which means that the option was again ITM from 8/24 to 9/8.
```{r}
df_market_history %>% 
  filter(underlying == "SPY") %>% 
  filter(expiration == as.Date("2013-09-21")) %>% 
  filter(type == "put") %>% 
  filter(strike == 166) %>% 
  arrange(trade_date)
```



12. This query shows all the entries related to the September 2013 SPY 166 put in `df_option_history`.  Notice that after 8/23, all trade dates are represented. (Bonus Challenge:  use the `bizdays` package to verify that this is the case - this will probably be time consuming to figure out so don't worry about it too much unless you're feeling like a gangster.)

```{r}
df_option_history %>% 
  filter(underlying == "SPY") %>% 
  filter(expiration == as.Date("2013-09-21")) %>% 
  filter(type == "put") %>% 
  filter(strike == 166) %>% 
  arrange(trade_date)


```


**Summary**

Here are some facts about `df_option_history`.

1. Once an option shows up in `df_market_history` it will show up in `df_option_history` starting on that trade date, and *for all subsequent trade dates until expiration*.


2. When we do backtesting, if our trading strategy requires trading on a given `trade_date`, we will trade only OTM options, hece the options we will trade must be in `df_market_history` on that date.

3. We will then use `df_option_history` to calculate the daily pnl for that trade, for all `trade_dates` on which we hold the posiiton.














