---
title: "Q & A"
output: html_document
---

If a question comes up that I think would be worth answering for the entire class then I will post the answer here.

&nbsp;

#### 9/5/2018
When working with dataframes, the `$` notation can be combined with the `[]` notation.  In order for this code to run, `data_options_intro.csv` needs to be in your working directory.

```{r message=FALSE, include=TRUE}
library(readr)
df_options <- read_csv("data_options_intro.csv")
```

```{r}
df_options$strike[3:6]
```

Describe in word what this code is doing relative to `df_options_data`.

&nbsp;

#### 9/4/2018
There was a piece of bad code in the subsetting tutorial.

```{r}
##> df_options$(strike:implied_vol)
```

If you run this code you will get an error that reads:



<center>
`Error: unexpected '(' in "df_options$("`
</center>
\ 

I thought that there was a way to use the `$` to select a range of columns, but that's not the case.  The `$` notation is intended to select one column at a time.

There are ways to select multiple columns of a dataframe in `base` R, but they are all messy and I don't recommend them.  If you want to select multiple columns of a dataframe, use the `select()` function of the `dplyr` package.
